---
title: "GBSG2 Exploration"
author: "Sep Dadsetan"
date: "4/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(skimr)
library(naniar)
library(GGally)
library(rpart) # recursive partitioning (decision tree)
library(caret)
```

## Load Data

```{r}
data("GBSG2", package = "TH.data")
gbsg2 <- GBSG2 %>% as_tibble()
```

### Feature Descriptions

- **horTh**: hormonal therapy, a factor at two levels no and yes.
- **age**: of the patients in years.
- **menostat**: menopausal status, a factor at two levels `pre` (premenopausal) and `post` (postmenopausal).
- **tsize**: tumor size (in mm).
- **tgrade**: tumor grade, a ordered factor at levels `I < II < III`.
- **pnodes**: number of positive nodes.
- **progrec**: progesterone receptor (in fmol).
- **estrec**: estrogen receptor (in fmol).
- **time**: recurrence free survival time (in days).
- **cens**: censoring indicator (0- censored, 1- event)

## Explore Characteristics

Take a look at characteristics like dimension, features, missingness, distribution, etc

```{r}
# Print tibble so we can look at the col types, values, types
gbsg2

# Use skim to look at summaries
skim(gbsg2)

# are there any missing values?
miss_var_summary(gbsg2)
```

## Manipulate Data

Make any transformations or imputations necessary to create a tidy dataset.

- Is the data wide/long?
- Are columns appropriately type'd?

## Visualize broadly

```{r}
ggpairs(gbsg2, mapping = aes(color = menostat))

library(survival)
gbsg_surv <- survival::survfit(Surv(time, cens) ~ tgrade, data = gbsg2)
ggsurv(gbsg_surv)
```

```{r}
# Break up sample into train, test, validation
set.seed(123)
assignment <- sample(1:3, nrow(gbsg2), prob = c(0.7, 0.15, 0.15), replace = T)

# add groups to gbsg2 tibble
gbsg2 <- gbsg2 %>% mutate(group = assignment)

mod <- rpart(formula = cens ~ ., 
      data = gbsg2, 
      method = 'class',
      parms = list(split = "gini"))

pred <- predict(object = mod, 
        newdata = gbsg2_test, 
        type = 'class')

confusionMatrix(data = pred, 
                reference = as.factor(gbsg2_test$cens))
```